### 2021-09-10 paper/localpsf.tex
optimal weighting function proof complete, except for lemma reference

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = v_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $v_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
\begin{equation}
\label{eq:optimal_v}
v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
\end{equation}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{v} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{v} \in \mathbb{R}^r$ is the vector with $i$th component $v_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation}
\label{eq:S_matrix}
S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
\end{equation}
\end{enumerate}
In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.

Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, then post-processing these functions. This is shown in Algorithm \ref{alg:weighting_functions_incremental}. The Poisson PDE solves, while the most expensive part of this procedure, can be performed cheaply with multigrid.

@@

\begin{proof}
	Let
	\begin{equation*}
		X := \{v \in H^2_N(\Omega) : v(x_i) = v_i \text{ for }i=1,\dots,r\}
	\end{equation*}
	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
	\begin{equation*}
	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,r\}
	\end{equation*}	
	denote the tangent space to $X$. Note that pointwise evaluation of functions in $H^2_N(\Omega)$ is well-defined because the Sobolev impedding theorem implies that $H^2_N(\Omega)$ is continuously imbedded in the spacde of continuous functions $C^0(\Omega)$ whenever the spatial dimension satisfies $d < 4$, and here we consider $d \in \{1,2,3\}$. Hence the optimization problem, the feasible set $X$ and the tangent space $X_0$ are well-defined.
\end{proof}

>>>>

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = f_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $f_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
\begin{equation}
\label{eq:optimal_v}
v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
\end{equation}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{f} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{f} \in \mathbb{R}^r$ is the vector with $i$th component $f_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation}
\label{eq:S_matrix}
S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
\end{equation}
\end{enumerate}
In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.

Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, then post-processing these functions. This is shown in Algorithm \ref{alg:weighting_functions_incremental}. The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.

@@

\begin{proof}
	Let
	\begin{equation*}
		X := \{v \in H^2_N(\Omega) : v(x_i) = v_i \text{ for }i=1,\dots,r\}
	\end{equation*}
	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
	\begin{equation*}
	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,r\}
	\end{equation*}	
	denote the tangent space to $X$. Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$. The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \ref{eq:wi_optimization_problem_generic} is well-defined.

	The proof now proceeds in three steps. First, we show that $v_*$ is feasible ($v_* \in X$). Second, we show that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} if it satisfies a certain variational equation. Third, we verify that $v_*$ satisfies this variational equation.

	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = f_i$ for $i=1,\dots,r$. We have
	\begin{align*}
	v_*(x_i) &= \beta + \sum_{j=1}^r c_j \theta_j(x_i) \\
	&= \beta + \sum_{j=1}^r c_j \left(\psi_i, \Delta \theta_j\right)_{L^2(\Omega)} + \avg\left(\theta_j\right) \\
	&= \beta + \sum_{j=1}^r c_j \left(\psi_i, \psi_j\right)_{L^2(\Omega)} \\
	&= \beta + \sum_{j=1}^r c_j S_{ij}
	\end{align*}
	In the first line we used the definition of $v_*$ in \eqref{eq:optimal_v}. In the second line we used Lemma LEM. In the third line we used the facts that, by construction, $\Delta \theta_j = \psi_j$ and $\int_\Omega \theta_j(x) dx = 0$. In the fourth line we used the definition of $S$ in \eqref{eq:S_matrix}. Hence the condition $v_*(x_i) = f_i$ for $i=1,\dots,r$ may be written in matrix form as
	\begin{equation*}
	S \mathbf{c} + \beta \mathbf{1} = \mathbf{f},
	\end{equation*}
	and this must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\mathbf{c}$ and $\beta$. So $v_* \in X$.

	\paragraph{Optimality} Since $X$ is an affine subspace, and $X_0$ is the tangent space to $X$, any $v \in X$ may be written in the form $v = v_* + p$ for some $p \in X_0$. We may therefore reframe optimization problem \ref{eq:wi_optimization_problem_generic} as 
	\begin{equation}
	\label{eq:p_optimization}
		\min_{p \in X_0} J(p),
	\end{equation}
	where
	\begin{equation*}
		J(p) := \frac{1}{2}||\Delta (v_* + p)||^2.
	\end{equation*}
	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} is equivalent to showing that $0$ is the unique minimizer of \eqref{eq:p_optimization}.

	By direct expansion, we have
	\begin{equation}
		\label{eq:Ju_expanded}
		J(p) = J(0) + \left(\Delta u, \Delta p\right)_{L^2(\Omega)} + \frac{1}{2}||\Delta p||^2.
	\end{equation}
	Since $p$ has Neumann zero boundary conditions, if $\Delta p = 0$, then $p$ is constant on each connected component of $\Omega$. Furthermore, since $p(x_i)$ is zero for at least one point $x_i$ per connected component of $\Omega$, if $\Delta p = 0$ then $p=0$. Hence, by contrapositive we have
	\begin{align*}
	\label{eq:contrapositive_laplacian_zero}
		p \neq 0 \quad \implies \quad& 0 < ||\Delta p||_{L^2(\Omega)}\\
		\implies \quad& J(0) < J(p) + \left(\Delta v_*, \Delta p\right)_{L^2(\Omega)}.
	\end{align*}
	If $v_*$ satisfies the following variational equation,
	\begin{equation}
	\label{eq:variational_up}
		(\Delta v_*, \Delta p)_{L^2(\Omega)} = 0 \quad \text{for all~} p \in X_0,
	\end{equation} 
	then
	\begin{equation*}
		p \neq 0 \quad \implies \quad J(0) < J(p).
	\end{equation*}
	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} therefore reduces to showing that $v_*$ satisfies variational equation \eqref{eq:variational_up}.

	\paragraph{Variational equation} To show that $v_*$ satiafies variational equation \eqref{eq:variational_up}, notice that
	\begin{align*}
	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \theta_i, \Delta p)_{L^2(\Omega)} \\
	=& \sum_{i=1}^r c_i \left(p(x_i) + \avg\left(p\right)\right) \\
	=& \avg\left(p\right) \sum_{i=1}^r c_i. 
	\end{align*}
	In the first line we used the definition of $\theta_i$ and the fact that the Laplacian of a constant is zero. In the second line we used Lemma LEM. In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$. Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^r c_i = 0$, which may be written in matrix form as $\mathbf{1}^T \mathbf{c} = 0$. This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
\end{proof}
